---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# odds

<!-- badges: start -->
<!-- badges: end -->

The goal of `{odds}` is to provide an on-disk data-storage of native R object, for cross-session data access.

## Installation

You can install the dev version of `{odds}` from GitHub with:

``` r
remotes::install_github("colinfay/odds")
```

## Basic use

The main goal of `{odds}` is to create a data storage architecture, on disk, so that you can access the values from one session to another.

### How it works

By default, the storage is done at `~/.kbk`, but it can be changed when creating the storage object.

__Note that the path is passed through `fs::path_norm()`, which doesn't treat `~` the same way as base R on Windows.__

```{r}
library(odds)
st <- Storage$new()
```

There are two main methods: `set()` and `get()`. 
The first saves a value under a name on the disk, the second retrieve this value from the storage. 

```{r}
st$set(head(iris), "a")
st$get("a")
```

Storages can be namespaced, and the default is "global",

```{r}
nsp <- paste(sample(letters, 3), collapse = "")

st$set(mtcars, "a", namespace = nsp)
st$get("a", namespace = nsp)
```

### Cross session access

Let's create an object in another R session: 

```{r}
library(callr)
rx <- r_bg(
  function(){
    library(odds)
    st <- Storage$new()
    st$set(head(airquality), "ping", namespace = "blop")
  }
)
rx$wait()
```

It's now accessible in the first session:

```{r}
st$get("ping", namespace = "blop")
```

Namespaces can be deleted: 

```{r}
st$remove_namespace("blop")
```


## Overhead

Of course, reading from disk adds some overhead, but for small to medium size objects, the cost of `get`ting from disk instead of reading for RAM is pretty small.

```{r}
library(dplyr)
library(ggplot2)
st$set(diamonds, "dm", "bench")
bench::mark(
  ram = {
   diamonds %>% filter(cut == "Ideal")
  }, 
  disk = {
    st$get("dm", "bench") %>% 
      filter(cut == "Ideal")
  }
)
```

`set` and `get` are powered by `{qs}` `qread()` and `qwrite()` and take the same arguments, so you can use parameters to these functions to speed up the read and write timing. 

Read the `{qs}` benchmark [online](https://github.com/traversc/qs#summary-table).

## Acknowledgment 

This package heavily relies on the `{qs}` package. 
Thanks to the package authors for their work.
